Tutorial: Quantifying Spatially Resolved Hydration Thermodynamics Using Grid Inhomogeneous Solvation Theory [Article v\versionnumber]

### 2024-07-01 manuscript/manuscript.tex
Add missing code in 4.3.5, fix code indent

Information about the grid is stored in the grid property of the gist object, structural data is stored as an \software{mdtraj} trajectory in the struct property:
%TODO we are missing some code here
Gisttools also provides OpenDX writing capabilities to visualize the calculated or modified GIST data. For visualization, you'll most likely want to write out the density-normalized GIST quantities (X\_dens):

@@

difference = compl_rdf - biotin_rdf - strept_rdf
ax2.plot(bins, np.cumsum(difference))
ax2.axvline(cutoff, color='k', linestyle='--')
ax2.set_xlabel('distance to biotin [Å]')
ax2.set_ylabel('$\Delta \Delta G$ [kcal/mol]')
ax2.grid()
plt.show()
\end{lstlisting}
The expected result is shown in Figure~\ref{fig-dg-sums}.
\begin{figure}[h]
	\centering
	\includegraphics[width=1.0\linewidth]{figures/deltaG-difference.png}
	\caption{Left: $\dgsolv$ contributions around the location of biotin in the binding pocket, evaluated in the biotin, streptavidin, and complex systems. Each line represents a cumulative sum plotted against the distance to biotin. Right: The hydration contribution to binding, evaluated as the difference between the lines in the left panel. The vertical dashed line is at \SI{12}{\angstrom} and represents the chosen integration cutoff.}\label{fig-dg-sums}
\end{figure}
Compute the energy (\inlinecode{Eall}) and entropy (\inlinecode{dTSsix}) contributions separately.
 To make this easier, gisttool includes a function to integrate in a radius around atom centers:
\begin{lstlisting}[style=python]
import pandas as pd
centers = {'biotin': biotin_x, 
	'strept': strept_x, 
	'compl': compl_x}
gist_objs = {'biotin': biotin, 
	'strept': strept, 
	'compl': compl}
cols = ['Eall_dens','dTSsix_dens']
rmax = 12
results = {
	name: gist.integrate_around(
	cols, rmax, centers[name])
	for name, gist in gist_objs.items()}
print(pd.DataFrame(results))
\end{lstlisting}
%TODO: This has a higher indent than the other python snippets in the pdf, somehow. Unsure why...
You will find that the energy disfavors binding, because we do not yet include the interaction energy between biotin and streptavidin.
Using the \inlinecode{energy} command in \software{cpptraj}, you can compute this interaction energy.
We recommend using PME in combination with PME-GIST, but not with GPU-GIST to be in line with the different ways the energy is calculated in the two GIST methods.
An example \software{cpptraj} input might look like:
\begin{lstlisting}[style=cpptraj]
parm solvated.parm7
trajin md-01.nc 1 last 100
energy complex ^1,2 # etype pme
energy strept ^1 # etype pme
energy biotin ^2 # etype pme
go
diff = complex[total] - strept[total] - biotin[total]
writedata energy.dat diff complex[total] \
strept[total] biotin[total]
avg(complex[total])
avg(strept[total])
avg(biotin[total])
avg(diff)
\end{lstlisting}
It has been shown \cite{Chen2021,Waibl2022-gist-solvents} that the solvation entropy in water is best approximated by 0.6 times the first order entropy.
Also compute the scaled entropy and check the effect on the binding affinity.
The expected results are summarized in Table~\ref{tab_dg_monomers_dimer}.

>>>>

Information about the grid is stored in the grid property of the gist object, structural data is stored as an \software{mdtraj} trajectory in the struct property:
\begin{lstlisting}[style=python]
gist.grid # Grid object
gist.struct # mdtraj.Trajectory object
\end{lstlisting}
Gisttools also provides OpenDX writing capabilities to visualize the calculated or modified GIST data. For visualization, you'll most likely want to write out the density-normalized GIST quantities (X\_dens):

@@

difference = compl_rdf - biotin_rdf - strept_rdf
ax2.plot(bins, np.cumsum(difference))
ax2.axvline(cutoff, color='k', linestyle='--')
ax2.set_xlabel('distance to biotin [Å]')
ax2.set_ylabel('$\Delta \Delta G$ [kcal/mol]')
ax2.grid()
plt.show()
\end{lstlisting}
The expected result is shown in Figure~\ref{fig-dg-sums}.
\begin{figure}[h]
	\centering
	\includegraphics[width=1.0\linewidth]{figures/deltaG-difference.png}
	\caption{Left: $\dgsolv$ contributions around the location of biotin in the binding pocket, evaluated in the biotin, streptavidin, and complex systems. Each line represents a cumulative sum plotted against the distance to biotin. Right: The hydration contribution to binding, evaluated as the difference between the lines in the left panel. The vertical dashed line is at \SI{12}{\angstrom} and represents the chosen integration cutoff.}\label{fig-dg-sums}
\end{figure}
Compute the energy (\inlinecode{Eall}) and entropy (\inlinecode{dTSsix}) contributions separately.
 To make this easier, gisttool includes a function to integrate in a radius around atom centers:
\begin{lstlisting}[style=python]
import pandas as pd
centers = {'biotin': biotin_x, 
    'strept': strept_x, 
    'compl': compl_x}
gist_objs = {'biotin': biotin, 
    'strept': strept, 
    'compl': compl}
cols = ['Eall_dens','dTSsix_dens']
rmax = 12
results = {
    name: gist.integrate_around(
    cols, rmax, centers[name])
    for name, gist in gist_objs.items()}
print(pd.DataFrame(results))
\end{lstlisting}
You will find that the energy disfavors binding, because we do not yet include the interaction energy between biotin and streptavidin.
Using the \inlinecode{energy} command in \software{cpptraj}, you can compute this interaction energy.
We recommend using PME in combination with PME-GIST, but not with GPU-GIST to be in line with the different ways the energy is calculated in the two GIST methods.
An example \software{cpptraj} input might look like:
\begin{lstlisting}[style=cpptraj]
parm solvated.parm7
trajin md-01.nc 1 last 100
energy complex ^1,2 # etype pme
energy strept ^1 # etype pme
energy biotin ^2 # etype pme
go
diff = complex[total] - strept[total] - biotin[total]
writedata energy.dat diff complex[total] \
strept[total] biotin[total]
avg(complex[total])
avg(strept[total])
avg(biotin[total])
avg(diff)
\end{lstlisting}
It has been shown \cite{Chen2021,Waibl2022-gist-solvents} that the solvation entropy in water is best approximated by 0.6 times the first order entropy.
Also compute the scaled entropy and check the effect on the binding affinity.
The expected results are summarized in Table~\ref{tab_dg_monomers_dimer}.
